---
title: "Solar Radiation EDA & prediction"
author: "Jonathan Bouchet"
date: "`r Sys.Date()`"
output:
 html_document:
    fig_width: 10
    fig_height: 7
    toc: yes
    number_sections : yes
    code_folding: show
---

```{r setup}
#options(width=100)
#knitr::opts_chunk$set(out.width='1000px',dpi=200,message=FALSE,warning=FALSE)
knitr::opts_chunk$set(message=FALSE,warning=FALSE)
```

```{r}
#load packages and csv file
library(ggplot2)
library(dplyr)
library(gridExtra)
library(Amelia)
library(corrplot)
library(ggfortify)
library(RColorBrewer)
library(forecast)
```

#EDA
```{r}
df<-read.csv("SolarPrediction.csv",sep = ",", dec = ".", header = TRUE)
any(is.na(df))
str(df)
```

We need :

* to convert the `Data` feature to extract `year`,`month`,`day` and `hour` information
* to calculate the duration of the day, as the difference in time between `TimeSunSet` and `TimeSunRise`

##Utils
```{r}
getDate<-function(x,pos1,pos2){
  if(pos1==1){
  val<-as.numeric(strsplit(strsplit(as.character(x)," ")[[1]][pos1],'/')[[1]][pos2])
  }
  else if(pos1==3 & pos2==0){
    val<-as.factor(strsplit(strsplit(as.character(x)," ")[[1]][pos1],'/')[[1]])
  }
  return(val)
}

getTIME<-function(x,pos){
  val<-strsplit(as.character(x),":")[[1]][pos]
  return(as.numeric(val))
}
```

```{r}
df$Month<-sapply(df$Data,getDate,1,1)
df$Day<-sapply(df$Data,getDate,1,2)
df$Year<-sapply(df$Data,getDate,1,3)
df$TimeAbbr<-sapply(df$Data,getDate,3,0)
df$hour<-sapply(df$Time,getTIME,1)
df$minute<-sapply(df$Time,getTIME,2)
df$sec<-sapply(df$Time,getTIME,3)
head(df)
```

```{r}
mymonths<-c("January","February","March","April","May","June","July","August","September","October","November","December")
df$MonthAbb <- mymonths[ df$Month ]
df$ordered_month <- factor(df$MonthAbb, levels = month.name)
#posixct for time serie
df$DateTs<-as.POSIXct(paste0(df$Year,'-',df$Month,'-',df$Day,' ',as.character(df$Time)), format="%Y-%m-%d %H:%M:%S")
df$DailyTs<-as.POSIXct(as.character(df$Time), format="%H:%M:%S")
df$DiffTime<-as.numeric(difftime(
  as.POSIXct(paste0(df$Year,'-',df$Month,'-',df$Day,' ',as.character(df$TimeSunSet)), format="%Y-%m-%d %H:%M:%S"),
  as.POSIXct(paste0(df$Year,'-',df$Month,'-',df$Day,' ',as.character(df$TimeSunRise)), format="%Y-%m-%d %H:%M:%S"),
  units='sec'))
```

##Radiation level
```{r}
ggplot(data=df,aes(x=Radiation,fill=ordered_month)) + 
  geom_histogram(bins=100) + 
  scale_y_log10() + 
  scale_fill_manual(name="",values=c("#46ACC8","#F21A00","#EBCC2A","#0B775E")) + 
  theme(legend.position='top') + 
  facet_wrap(~ordered_month) +
  xlab('Radiation level [W/m^-2]') + ylab('Count')
```

Comments :

* we see a large peak for low Radiation (~0, meaning no Radiation) then smaller activity
* September is the month with a more pronounce peak at higher Radiation (>1000W/m^-2)

```{r}
ggplot(data=df,aes(x=DiffTime,y=Radiation)) + 
  geom_point(aes(color=ordered_month)) +
  scale_color_manual(name="",values=c("#46ACC8","#F21A00","#EBCC2A","#0B775E")) + 
  theme(legend.position='top') + xlab("SunSet -SunRise [sec]")
```

* This is an expected result : since the day is longer in September vs. all the other month, so the difference between sunrise and the sunset is longer
* it also seems that Radiation with higher level appear more often in September (or might just be a statistical issue, day longer so more probability to get a high Radiation)

##Mean Radiation per Day
```{r}
df %>% select(ordered_month,Day,Radiation) %>% 
  group_by(ordered_month,Day) %>% 
  summarise(dailyRad = mean(Radiation)) %>% 
  ggplot(aes(x=ordered_month,y=dailyRad,color=dailyRad)) + 
  scale_color_gradientn(colours=rev(brewer.pal(10,'Spectral'))) + 
  geom_boxplot(colour='black',size=.4,alpha=.5) + 
  geom_jitter(shape=16,width=.2,size=2) +
  xlab('') + ylab('') + theme(legend.position='top')
```

###Radiation for a given day (as a test plot)
```{r,eval=F}
ggplot(data=filter(df,Day==1 & ordered_month=='September'),aes(x=DateTs,y=Radiation)) + 
  geom_line() + xlab('')
df %>% filter(Day==1 & ordered_month=='September') %>% summarise(mean(Radiation))
```

##Time Serie
```{r}
dailyRadiation<-as.data.frame(df %>% 
                                select(Month,Day,Radiation) %>% 
                                group_by(Month,Day) %>% 
                                summarise(dailyRad = mean(Radiation)))
#(re)create a Date format
dailyRadiation$dateTS<-as.Date(paste0('2016-',dailyRadiation$Month,'-',dailyRadiation$Day))
#weekly moving average
dailyRadiation$ma_week<-ma(dailyRadiation$dailyRad,order=7)
dailyRadiation$ma_2weeks<-ma(dailyRadiation$dailyRad,order=14)

ggplot() + 
  geom_line(data=dailyRadiation,aes(x=dateTS,y=dailyRad,color='by_day'),alpha=.75) + 
  geom_line(data=dailyRadiation,aes(x=dateTS,y=ma_week,color='by_week'),size=2) + 
  geom_line(data=dailyRadiation,aes(x=dateTS,y=ma_2weeks,color='by_2weeks'),size=2) + 
  xlab('') + ylab('Average Solar Radiation') + 
  scale_colour_manual(name="",values=c(by_day="black",by_week="#46ACC8",by_2weeks="#F21A00")) + 
  theme(legend.position='top')

```

* We may not have enough data to make a time serie analysis
* it seems (with this data) that we cannot identify any periodicity

#Prediction
to do
<hr>
<strong>History :</strong>

* _version 1 : initial commit_ 

<hr>